Sherlock topic model paper

### 2018-05-15 main.tex
saving progress half way through the event segmentation results section

\subsubsection{Projecting verbal recall into stimulus space}
After watching the episode, participants verbally recalled (in order) as much of the episode as they could.  We used the same topic model (fit with the text descriptions of the video) to transform participants' verbal recall transcripts (split by sentence in overlapping chunks of 10 sentences). The result was a sentences (range: X to X) by topics (100) matrix for each participant, where each row represented the estimated mixture of topics for a given window of sentences during recall. In order to average the recall models across participants, we resampled each model at the resolution of the video model (1976 timepoints). The individual models can be seen in Supp Fig X and the group average is plotted in \ref{fig:model}c. Note that the topics were derived solely from the text descriptions of the stimulus, and so the topic models estimated for recall are entirely dependent on the topics in the episode. In effect, this approach projects the recall transcripts into the video 'stimulus space', allowing us to quantify the dynamics of verbally reported memories as well as relate them to the experiences from which they were born.

Next, we investigated the temporal structure of the recall matrices. For each participant, we computed a timepoint-by-timepoint correlation matrix from the recall model. Like the video model, the recall correlation matrices exhibited blocky structure along the diagonal, albeit .

@@

 However, the size/number of blocks varied across participants.  To quantify this structure, we estimated the number of recall `events' (k)  discovered in each participants' recall matrix (see Methods for details) using the event segmentation approach developed by Baldassano et al 2017.  Briefly, this algorithm implements a variant of a Hidden Markov Model to estimate a set of events (or states) given a multivariate time series and a k value (number of states). Figure X displays example correlation matrices for the participants with the lowest and highest k value.  We found that the estimated k value strongly correlated with the number of (hand-annotated) correctly remembered scenes across participants (Figure Xx, r=.67, p=.003). Thus, the blocky temporal structure of the recall matrices computed with our model is related to the variability in memory performance across participants.

Another interesting feature of the recall matrices is the off-diagonal structure, which represents the use of language that generalizes to more than one event. To characterize this aspect of the recall structure, we averaged the rows of the recall matrices according to the k events computed in the analysis described above.  This reduced the recall matrices from \# of sentences to k rows. Then, we computed k by k correlation matrices, representing the similarity of each event to each other event. To visualize these recall `networks', we embedded the k by topics matrices into a 2-dimensional space using Multidimensional Scaling (MDS), where each dot represents an event and the color of the dot represents the event index.  Then, we connected each event with a line where the color and line weight was proportional to the strength of the correlation between the events. These network plots can be seen in Fig X. To hypothesized that stronger correlations between recall events would

\subsubsection{Temporally aligning the stimulus and recall}
While most participants were able to recall most of the scenes in order, the rate of verbal description varied by participant and timepoint (Supp Fig X). Because of this participant-dependent non-linear relationship between stimulus and recall time, averaging the video-recall correlation plots across participants attenuates the robust within-participant effects. While further exploration of individual differences in recall rates is interesting, our primary goal here was to develop a model capable of accurately characterizing the content recovered during recall and explore how that varied across participants.  To that end, we leveraged dynamic time warping (DTW, REF) to temporally align the stimulus and recall matrices.  The algorithm returns a path of indices that bring the two timeseries into maximal temporal alignment (within constraints, see Methods for details).  For each participant, we computed the path and then warped the video model and recall model according to it. The length of the warp path varied by participant, and so the length (number of timepoints) of the warped video/recall matrices also varied by participant (range: X-X). To account for these differences, we resampled the timeseries back to the length of the stimulus (1976 time samples).

@@

% \subsubsection{Accounting for nonlinearities between video and recall models}
% For many participants, the rate of recall changed over time (see Supp Fig X for an example). This becomes problematic if averaging the video/recall correlation matrices because the different recall rates distort the group average matrix. To correct for differences in recall rates across participants, we used dynamic time warping (DTW).  Briefly, DTW finds the path that maximizes the temporal alignment between two timeseries (REF). For each participant, we used DTW to temporally align the video model and the participant-specific recall model. Because the path is different for each participant, the length of the warped matrices varied across participants (range: X to X). To account for this, we resampled the matrices back to their original shape (1976 x 100).

@@

% \subsection{fMRI analyses}
% Seventeen participants watched the first 50 minutes of Episode 1 of BBC's Sherlock. The video was split into two parts of approximately equal length (946 and 1030 TRs). All data were preprocessed and transformed to 3mm MNI space as described in the paper. Data were zscored across time at every voxel. 6mm smoothing was applied.
% Files are cropped so that all video-viewing data are aligned across participants, and all recall data are aligned to the scene timestamps below. The cropping includes a constant 3-TR (4.5 sec) shift to correct for hemodynamic lag.
%
% \subsubsection{Univariate subsequent memory analysis}
% To characterize brain regions involved in memory encoding, we conducted an analysis inspired by prior studies investigating the  `subsequent memory effect' (Paller and Wagner).  For each participant, we computed the correlation between topic vectors at each timepoint of the video model and recall model (they were temporally aligned with DTW and resampled to 1976 timesamples). The result was a 1 x 1976 vector of correlation values representing how similar each moment of the video model was to each moment of the recall model. Then, to compute the average correlation, we Fisher's z transformed the correlation vectors, averaged them together, and then inverse Fisher's z transformed them.  Finally, we mean centered the vector.  This 1 x 1976 vector represents memorable/unmemorable parts of the video, and be thought of as a `continuous' extension of a subsequent memory regressor.  We convolved the timeseries with a double gamma function (PARAMS) to account for the hemodynamic response, and used it as a model for each voxel's activity timeseries (using FSL software).  Finally, to aggregate across participants, we performed a mixed effects (ordinary least squares) analysis, thresholding at p<.001 and cluster correcting a p<.05.
%
% \subsubsection{Multivariate searchlight analyses}
% Our multivariate analyses were designed to capture brain regions whose timepoint-by-timepoint correlational structure mirrors the correlational structure of the video and recall topic models during video viewing. To that end, we conducted a searchlight analysis (5x5x5 voxel cube) where for each cube, we correlated the model timepoint-by-timepoint correlation matrix with the neural correlation matrix. To aggregate across participants, we computed Fisher's z transformed the correlations and then averaged.  To assess significance, we recomputed this group analysis 100 times, but randomly phase shifted the model (by \# of timepoints - 1) by the same amount for each participant but different amounts for each permutation to build a null distribution of correlation values. Finally, we thresholded the group averaged correlation maps where the `real' correlation value for a given voxel exceeded the 95th percentile of the null distribution.
%
% We first performed this analysis using the unwarped video model. Thus, the same model was applied to each participant and each searchlight cube. We then performed this analysis on the participant-wise recall models. To correct for non-linearities between the video model and recall models, for each participant we used DTW to temporally align the matrices. The algorithm recovers a path of coordinates that would bring the video and recall model in maximal temporal alignment. We used this path to warp the fMRI data and the recall model into alignment (separately for each participant). Then, we performed the exact same analysis as described above, using a searchlight to correlate the timepoint-by-timepoint recall and neural correlation matrices. The statistics for both analyses were computed as described above.

>>>>

\subsubsection{Projecting verbal recall into stimulus space}
After watching the episode, participants verbally recalled (in order) as much of the episode as they could.  We used the same topic model (fit with the text descriptions of the video) to transform participants' verbal recall transcripts (split by sentence in overlapping chunks of 10 sentences). The result was a sentences (range: X to X) by topics (100) matrix for each participant, where each row represented the estimated mixture of topics for a given window of sentences during recall. In order to average the recall models across participants, we resampled each model at the resolution of the video model (1976 timepoints). The individual models can be seen in Supp Fig X and the group average is plotted in \ref{fig:model}c. Note that the topics were derived solely from the text descriptions of the stimulus, and so the topic models estimated for recall are entirely dependent on the topics in the episode. In effect, this approach projects the participants' verbal recall into the video 'stimulus space', allowing us to quantify the dynamics of verbally reported memories as well as relate them to the experiences from which they were born.

Next, we investigated the temporal structure of the recall matrices. For each participant, we computed a timepoint-by-timepoint correlation matrix from recall models. Like the video model, the recall correlation matrices exhibited blocky structure along the diagonal (Supp Fig X). To visualize the group average, we resampled all of the recall models to the duration of the video model (1976 timepoints), computed the timepoint-by-timepoint correlation matrices and then averaged across partcipants.  Like the video correlation matrix, the average recall matrix was blocky along the diagonal, albeit noisier (\ref{fig:model}d).  Presumably this noise stems from the fact that there was variability in the amount of time each participant spent describing each event. We will come back to this point later in the paper.

\subsubsection{Correlating the video and recall models}
The participants in this study were instructed to recall as much of the episode as they could in the order it happened.  Thus, we expected that to the extent that participants followed these instructions and had reasonably good memory accuracy, the recall models should resemble the video model. In our first approach to test this hypothesis, we correlated each moment of the video model with each moment of each participants' resampled recall model. We then averaged these matrices together to get a single group averaged matrix.  The result was a viewing time (1976) by recall time (1976) correlation matrix representing the average relationship between every moment of the movie and recall models (\ref{fig:model}e). Notably, we observe significant correlation values (p<.05, permutation test described in \ref{sec:methods}) primarily along the diagonal, suggesting that our model captures the fact that participants recalled the episode in order. We also visualized this effect by projecting the matrices into a 3D space (reduced using Spectral Embedding [cite]) where the proximity of the trajectories at every timepoint represents how similar the topic vectors were. Visualizing the data in this way makes it clear that the overall shape and correlational structure of the video and average recall matrices is quite similar.

@@

\subsubsection{Segmenting the video and recall models into events}
One prominent feature of the video and recall correlation matrices (see \ref{fig:model}) is the strong blocky structure along the diagonal.  We hypothesized that this structure arises from transient stability in the language used to describe each scene of the video. If that were true, we reasoned that it would be possible to decode which particular scene a participant is describing by comparing the average topic model for a particular recall `event' with the average topic vector for each video event. The video event that is most similar to the given recall event is the event the participant is most likely describing.  To test this idea it is first necessary to systematically parse the matrices into events.  To do this, we leveraged the event segmentation approach developed by Baldassano et al 2017.  Briefly, this algorithm implements a variant of a Hidden Markov Model to estimate k events given a (multivariate) time series and a k value (number of states). We automatically determined a k value for the video model and each participant's recall model using the following equation:

% ADD EQUATION HERE

The algorithm determined 34 events for the movie model and a range of values (range=8-27, mean= 15.41, SD=5.6) for the recall models.  The events discovered for a representative participant can be seen in \ref{fig:eventseg}b. Importantly, the k values were highly correlated to hand annotated accuracy (r(16)=.67, p=.003) across participants. Thus, the blocky temporal structure in the recall matrices computed by our model is strongly related to the variability in memory performance across participants.

Next, we tested the hypothesis that the individual `blocks' along the diagonals of the recall matrices represent the recall of a particular video event. To this end, we created a video event model, which averaged together neighboring topic vectors that the were classified to be in the same event (\ref{fig:eventseg}c), resulting in an events (34) by topics (100) matrix.  We performed the same procedure for the recall matrices (see \ref{fig:eventseg}d for example). Then, we computed the correlation between these two event models, resulting in a video events (34) by recall events (8-27, depending on the participant) correlation matrix (see \ref{fig:eventseg}e). These matrices represent the extent to which each recall event is similar to each video event (for each participant). To determine which video events the recall events most likely referred to, for event recall event, we computed video event with the highest correlation (i.e. the argmax).  This is depicted in \ref{fig:eventseg}e as the cell highlighted with the black box for each column in the video-recall event correlation matrix.

Figure X displays example correlation matrices for the participants with the lowest and highest k value.  We found that the estimated k value strongly correlated with the number of (hand-annotated) correctly remembered scenes across participants (Figure Xx, r=.67, p=.003).

@@

@@
